---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![Codecov test coverage](https://codecov.io/gh/BarkleyBG/glmerGOF/branch/master/graph/badge.svg)](https://codecov.io/gh/BarkleyBG/glmerGOF?branch=master)
<!-- badges: end -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# glmerGOF

The goal of glmerGOF is to provide a goodness of fit test of the presumed Gaussian
distribution of the random effect in logistic mixed models fit with 
`lme4::glmer(family = "binomial")`. The method implemented is introduced in 
Tchetgen Tchetgen and Coull (2006):

Tchetgen Tchetgen, E. J., & Coull, B. A. (2006) _A Diagnostic Test for the Mixing Distribution in a Generalised Linear Mixed Model_. Biometrika, 93(4), 1003-1010. DOI: [10.1093/biomet/93.4.1003](https://doi.org/10.1093/biomet/93.4.1003)

## Installation

glmerGOF is available through github:

```{r, eval = FALSE}
remotes::install_github("BarkleyBG/glmerGOF")
```

## Usage

An introductory tutorial is provided that describes how to use this package. The function that implements the test is `glmerGOF::testGOF()`, which takes as mandatory input:

- a fitted `lme4::glmer()` model
- a fitted `survival::clogit()` model
- the original dataset
- a list providing two variable names

Once the two models are fitted, then test statistics can be found:

```{r, eval = FALSE}
TC_test <- testGOF(
  data = my_data,
  fitted_model_clogit = fit_clogit,
  fitted_model_glmm  = fit_glmm,
  var_names = list(DV = "y", grouping = "id"), 
  gradient_derivative_method = "simple"
)
```

The test results can be shown as:

```{r, eval = FALSE}
TC_test$results
# $D
# [1] 1.230103
# 
# $p_value
# [1] 0.267387
```


# Potential future developments

Please refer to the lifecycle badge for its current status. This package was created in early 2019 and may undergo future developments. Please email the maintainer if any of these changes are of interest, or if you would like to work on them:

- Ability to manually align glmer and clogit coefficients
- Parallelization backend
- Updated model options


# Contributor Code of Conduct
Please note that the 'glmerGOF' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

