---
title: "Introduction to glmerGOF package"
author: "Brian G. Barkley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an introduction and tutorial to the glmerGOF package. This package implements a test of the goodness of fit of the presumed Gaussian random effect distribution in the logistic mixed models fit with `lme4::glmer()`. The method is introduced in Tchetgen Tchetgen, E. J., & Coull, B. A. (2006) _A Diagnostic Test for the Mixing Distribution in a Generalised Linear Mixed Model_. Biometrika, 93(4), 1003-1010. \url{https://doi.org/10.1093/biomet/93.4.1003}. This test is implemented in `glmerGOF::testGOF()`, as described below.

# Using glmerGOF

We begin with a simple data example, with data simulated as shown in the [geex package](https://bsaul.github.io/geex/articles/articles/lme4_discussion.html#setup)

```{r}
library(glmerGOF)
set.seed(1)
#generate data; 50 clusters
n <- 50
m <- 4
beta <- c(2, 2)
id <- rep(1:n, each=m)
x <- rnorm(m*n)
b <- rep(rnorm(n), each=m)
y <- rbinom(m*n, 1, plogis(cbind(1, x) %*% beta + b))
my_data <- data.frame(y,x,id) 

knitr::kable(head(my_data), digits = 2)
```

The `testGOF()` function in this package takes arguments for the original data, as well as two models. One model is fit with `lme4::glmer()`, and the other is a conditional logistic model fit with `survival::clogit()`.

## Fitting the two models

The two models are fit outside of the `glmerGOF()` package. The `testGOF()` function attempts to match the coefficients of the two models to produce the test results.

### Generalized logistic mixed model

```{r}
library(lme4)
fit_glmm <- lme4::glmer(
  formula = y ~ x + (1|id),
  family = "binomial",
  data = my_data
)
```

### Conditional logistic model

```{r}
library(survival)
fit_clogit <- survival::clogit(
  formula = y ~ x + strata(id),
  data = my_data,
  method = "exact"
)
```

## Executing the GOF test

Once the models are fit, there are only a few more steps. One step is to provide a list of two important variable names, as strings. `DV` is the dependent variable in the model, and `grouping` is the variable indicating the cluster membership.

```{r}
variable_names <- list(DV = "y", grouping = "id")
```

Lastly, choose whether you want 
```{r}
deriv_method <- 
  # "Richardson" ## for slower but more accurate computations
  "simple" ## for faster computation

```

Now, testGOF can be run by passing in the data, models, and variable names:

```{r}
TC_test <- testGOF(
  data = my_data,
  fitted_model_clogit = fit_clogit,
  fitted_model_glmm  = fit_glmm,
  var_names = variable_names, 
  gradient_derivative_method = deriv_method
)

```

The test statistic follows a chi-squared distribution with degrees of freedom equal to the number of parameters:

```{r}
TC_test$results
```








